Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2012-05-17 extended_catalog/paper/extended_search.tex
new draft v12.0

We also performed a Monte Carlo study that (a) shows that we obtain
an unbiased fit of an extended source with a complicated spatial shape so long as
the source is correctly modeled and (b) quantifies the biases 
introduced by incorrectly modeling its spatial structure.
To assess this, we simulated the spatially extended ring-type SNR W44.
We selected W44 because it is the most significant extended source detected by the LAT 
that has a non-radially symmetric photon distribution \citep{w44}.

@@

Figure~\ref{bias_w44sim} shows that there are no systematic biases in the
distribution of best fit (a) flux (b) spectral index (c) \rsixeight
(d) elliptical eccentricity (e) elliptical orientation angle and (f) inner
ring fraction when the simulated extended sources were fit with the
elliptical ring spatial model. For the elliptical spatial models, \rsixeight
is computed by numeric integration.
Figures~\ref{bias_w44sim}a and~\ref{bias_w44sim}b show that
no significant bias is introduced by modeling the source with the 
other extended spatial models, while modeling the source as point-like
results in $\sim10\%$ and $\sim5\%$ biases in the fit flux and index,
respectively.  
Furthermore, Figure~\ref{bias_w44sim}c shows that the \rsixeight estimate of
the extension is very mildly biased ($\sim10\%$) toward higher values
when inaccurate spatial models are used, and thus \rsixeight for any of
these spatial models can be used as a reasonable measurement of the true
68\% containment radius for the source.
}

@@

Because the two point-like source hypothesis has
three more parameters than the single extended source hypothesis (two spatial
parameters and two spectral parameters compared to one extension parameter),
we note that the $\tsext>\tsinc$ test is mathematically equivalent to testing that
$\aic_\text{ext} < \aic_\text{2pts} - 6$.
Thus our criterion based on the \ts value is equivalent to requesting
that the \aic-based empirical support for the two-point-like-source model be
``considerably less'' than for the extended source model, following the
classification by \cite{aic_stats_book}.
}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/bias_w44sim_color.eps}
    \else
    \plotone{mc_plots/bias_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_bias
    \caption{
    \newtext{The distribution of fit parameters
    for the Monte Carlo simulations of W44.
    The plots show the distribution of 
    best fit (a) flux (b) spectral index (c) 68\% containment
    radius (d) eccentricity (e) elliptical orientation angle
    and (f) inner ring fraction. The dashed vertical lines represent
    the simulated values of the parameters.
    }
    }\label{bias_w44sim}
\end{figure}

>>>>

We also performed a Monte Carlo study to show that incorrectly modeling the
spatial extension of an extended source does not substantially bias
the spectral fit of the source, although it does alter the value of the \ts.
To assess this, we simulated the spatially extended ring-type SNR W44.
We selected W44 because it is the most significant extended source detected by the LAT 
that has a non-radially symmetric photon distribution \citep{w44}.

@@

Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b clearly show that
no significant bias is introduced by modeling the source as extended
but with an inaccurate spatial model, while a point-like source modeling
results in a $\sim10\%$ and $\sim5\%$ bias in the fit flux and index,
respectively.  
Furthermore, Figure~\ref{bias_w44sim}c shows that the \rsixeight estimate of
the extension size is very mildly biased ($\sim10\%$) toward higher values
when inaccurate spatial models are used, and thus represent a reasonable
measurement of the true 68\% containment radius for the source.
For the elliptical spatial models, \rsixeight is computed by numeric integration.

@@

The two point-like sources hypothesis has three more parameters than
the single extended source hypothesis (two more spatial parameters and
two more spectral parameters compared to one extension parameter), so the
comparison $\aic_\text{ext} < \aic_\text{2pts}$  is formally equivalent to
$\tsext + 6 > \tsinc$.  Our criterion for accepting extension ($\tsext > \tsinc$) 
is thus equivalent to requesting that the AIC-based empirical
support for the two point-like sources model be ``considerably less''
than for the extended source model, following the classification by
\cite{aic_stats_book}.
}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/bias_w44sim_color.eps}
    \else
    \plotone{mc_plots/bias_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_bias
    \caption{
    \newtext{The distribution of fit parameters
    for the Monte Carlo simulations of W44.
    The plots show the distribution of 
    best fit (a) flux (b) spectral index and (c) 68\% containment
    radius. The dashed vertical lines represent
    the simulated values of the parameters.
    }
    }\label{bias_w44sim}
\end{figure}
